---
title: "MonGame"
author: "Daniel Halteh"
date: "1/11/2018"
output:
  pdf_document: default
  html_document: default
---

```{r, error=TRUE, include=FALSE}

# Establishing the game board

gameboard <- data.frame(space = 1:40, title = c("Go" , "Mediterranean Avenue" , "Community Chest" , "Baltic Avenue" , "Income Tax" , "Reading Railroad" , "Oriental Avenue" , "Chance" , "Vermont Avenue" , "Connecticut Avenue" , "Jail" , "St. Charles Place" , "Electric Company" , "States Avenue" , "Virginia Avenue" , "Pennsylvania Railroad" , "St. James Place" , "Community Chest" , "Tennessee Avenue" , "New York Avenue" , "Free Parking" , "Kentucky Avenue" , "Chance" , "Indiana Avenue" , "Illinois Avenue" , "B & O Railroad" , "Atlantic Avenue" , "Ventnor Avenue" , "Water Works" , "Marvin Gardens" , "Go to jail" , "Pacific Avenue" , "North Carolina Avenue" , "Community Chest" , "Pennsylvania Avenue" , "Short Line Railroad" , "Chance" , "Park Place" , "Luxury Tax" , "Boardwalk"))
chancedeck <- data.frame(index = 1:15, card = c("Advance to Go" , "Advance to Illinois Ave." , "Advance to St. Charles Place" , "Advance token to nearest Utility" , "Advance token to the nearest Railroad" , "Take a ride on the Reading Railroad" , "Take a walk on the Boardwalk" , "Go to Jail" , "Go Back 3 Spaces" , "Bank pays you dividend of $50" , "Get out of Jail Free" , "Make general repairs on all your property" , "Pay poor tax of $15" , "You have been elected Chairman of the Board" , "Your building loan matures"))
communitydeck <- data.frame(index = 1:16, card = c("Advance to Go" , "Go to Jail" , "Bank error in your favor. Collect $200" , "Doctor's fees Pay $50" , "From sale of stock you get $45" , "Get Out of Jail Free" , "Grand Opera Night Opening" , "Xmas Fund matures" , "Income tax refund" , "Life insurance matures. Collect $100" , "Pay hospital fees of $100" , "Pay school tax of $150" , "Receive for services $25" , "You are assessed for street repairs" , "You have won second prize in a beauty contest" , "You inherit $100"))
```


```{r}
# Dice --------------------------------------------------------------------

dice <- function(verbose=FALSE){
  faces <- sample(1:6, 2, replace=TRUE)
  if(faces[1] == faces[2]){
    doubles = TRUE
  }
  else {
    doubles = FALSE
  }
  movement = sum(faces)
  if(verbose){
    cat("Rolled:", faces[1], faces[2], "\n")
  }
  return(list(faces=faces, doubles=doubles, movement=movement))
}





# Player Reference Class --------------------------------------------------


# a **very basic** reference class for our players
player <- setRefClass("player", 
  fields = list(
    pos = "numeric",      # position on the board
    verbose = "logical",
    in_jail = "logical",
    n_jail = "numeric"
  ), 
  methods = list(
    move_n = function(n) {
      if(verbose) cat("Player at:", pos, "\n")
      if(verbose) cat(" Player moves:", n, "\n")
      pos <<- pos + n
# Landing on Go to Jail ---------------------------------------------------------------------
      if(pos == 31){
        pos <<- 11
        in_jail <<- TRUE
      }
      if(pos > 40) pos <<- pos - 40
      if(verbose) cat(" Player now at:", pos,"\n")
    },
    go_2_space_n = function(n){
      if(verbose) cat("Player at:", pos,". \n")
      pos <<- n
# Landing on Go to Jail ---------------------------------------------------------------------      
      if(pos == 31){
        pos <<- 11
      }
      if(verbose) cat(" Player now at:", pos,".\n")
    },
    inc_jail_count = function(){
      n_jail <<- n_jail + 1
    },
    reset_jail_count = function(){
      n_jail <<- 0
    },
    jailed = function(a){
      if(a == TRUE){
        in_jail <<- TRUE
      } else if (a == FALSE){
        in_jail <<- FALSE
      }
    }
  )
)



# Drawing a Chance Card --------------------------------------------------------------------
chance <- function(p, n){ # pass in player ID and player$pos
        chance_sample <- sample(chancedeck$index, 1)
        if (chance_sample == 1){
          p$go_2_space_n(1)
        } 
          else if (chance_sample == 2){
           p$go_2_space_n(25) 
        } 
          else if(chance_sample == 3){
           p$go_2_space_n(12) 
        } 
          else if(chance_sample == 4){
            if(n == 23){
              p$go_2_space_n(29) 
          } else {
              p$go_2_space_n(13)
          }
        } 
          else if(chance_sample == 5){
            if(n == 8){
              p$go_2_space_n(16) 
          } 
            else if(n == 23){
              p$go_2_space_n(26) 
          } 
            else p$go_2_space_n(6) 
        } 
          else if(chance_sample == 6){
           p$go_2_space_n(6)
        } 
          else if(chance_sample == 7){
           p$go_2_space_n(40) 
        } 
          else if(chance_sample == 8){
           p$go_2_space_n(11)
           p$jailed(TRUE)
        } 
          else if(chance_sample == 9){
           p$move_n(-3)
        } 
          else{
           p$go_2_space_n(n)
        }
}


# Drawing a Community Chest Card -------------------------------------------------------------------
chest <- function(p, n){ 
    chest_sample <- sample(communitydeck$index, 1)
    if(chest_sample == 1){
      p$go_2_space_n(1)
    } else if(chest_sample == 2){
      p$go_2_space_n(11)
      p$jailed(TRUE)
    } else
      p$go_2_space_n(n)
}




# Space Tracking Reference Class ------------------------------------------

# a *basic* reference class to keep track of where people landed
tracking <- setRefClass("tracking",
  fields = list(
    tally = "numeric"
  ),
  methods = list(
    increase_count = function(n){
      tally[n] <<- tally[n] + 1
    }
  )
)

space_tracking <- tracking$new(tally = rep(0,40))


# Jail Functionality -----------------------------------------------------------
   
jail <- function(p){
  roll <- dice()
  # If player rolls a double, he leaves jail
  if(roll$doubles == TRUE){
    p$jailed(FALSE)
    p$reset_jail_count()
    p$move_n(roll$movement)
  }
  else if (p$n_jail == 3){
    p$jailed(FALSE)
    p$reset_jail_count()
    p$move_n(roll$movement)
  } 
  else(p$go_2_space_n(11))
}




# Taking a turn -----------------------------------------------------------

taketurn <- function(p, track){
  #### If player is in jail ####
  if(p$in_jail == TRUE){
    p$inc_jail_count()
    jail(p)
    track$increase_count(p$pos)
  }
  
  #### If player not in jail ####
  else if(p$in_jail == FALSE){
    # First roll
    roll <- dice()
    # ROUTE 1: NOT A DOUBLE
    if(roll$doubles == FALSE){
      p$move_n(roll$movement)
      if(p$pos == 8 | p$pos == 23 | p$pos == 37){ # calls chance function if applicable
        chance(p, p$pos)
      }
      else if(p$pos == 3 | p$pos == 18 | p$pos == 34){ # calls chest function if applicable
        chest(p, p$pos)
      }
      track$increase_count(p$pos) # Tracks at the end of each turn
    }
    
    # ROUTE 2: DOUBLE
    else if(roll$doubles == TRUE){
      p$move_n(roll$movement)
      if(p$pos == 8 | p$pos == 23 | p$pos == 37){ # calls chance function if applicable
        chance(p, p$pos)
      }
      else if(p$pos == 3 | p$pos == 18 | p$pos == 34){ # calls chest function if applicable
        chest(p, p$pos)
      }
      track$increase_count(p$pos) # Tracks at the end of each turn
      
# Roll again for rolling doubles ---------------------------------------------------------------------      
      # Roll 2
      if(p$in_jail == FALSE){
      roll2 <- dice()
      # ROUTE 2.1: NOT A DOUBLE
      if(roll2$doubles == FALSE){
        p$move_n(roll2$movement)
        if(p$pos == 8 | p$pos == 23 | p$pos == 37){ # calls chance function if applicable
          chance(p, p$pos)
        }
        else if(p$pos == 3 | p$pos == 18 | p$pos == 34){ # calls chest function if applicable
          chest(p, p$pos)
        }
        track$increase_count(p$pos) # Tracks at the end of each turn
      }
      
      # ROUTE 2.2: DOUBLE
      else if(roll2$doubles == TRUE){
        p$move_n(roll2$movement)
        if(p$pos == 8 | p$pos == 23 | p$pos == 37){ # calls chance function if applicable
          chance(p, p$pos)
        }
        else if(p$pos == 3 | p$pos == 18 | p$pos == 34){ # calls chest function if applicable
          chest(p, p$pos)
        }
        track$increase_count(p$pos) # Tracks at the end of each turn
        
# Roll again (third roll) for rolling doubles ---------------------------------------------------------------------  
        # Roll 3
        if(p$in_jail == FALSE){
        roll3 <- dice()
        # ROUTE 2.2.1: NOT A DOUBLE
        if(roll3$doubles == FALSE){
          p$move_n(roll3$movement)
          if(p$pos == 8 | p$pos == 23 | p$pos == 37){ # calls chance function if applicable
            chance(p, p$pos)
          }
          else if(p$pos == 3 | p$pos == 18 | p$pos == 34){ # calls chest function if applicable
            chest(p, p$pos)
          }
          track$increase_count(p$pos) # Tracks at the end of each turn
        }
# Go to jail for rolling three doubles --------------------------------------------------------------------- 
        # ROUTE 2.2.2: DOUBLE (JAIL)
        else if(roll3$doubles == TRUE){
          p$go_2_space_n(11)
          p$jailed(TRUE)
          track$increase_count(p$pos) # Tracks at the end of each turn
          
        }
      }
      }

    } 
    
    }
  }
}



# Simulation --------------------------------------------------------------

player1 <- player$new(pos = 1, verbose = TRUE, in_jail = FALSE, n_jail = 0)  # create new players
player2 <- player$new(pos = 1, verbose = TRUE, in_jail = FALSE, n_jail = 0)




# Running the simulation --------------------------------------------------

set.seed(1)

space_tracking <- tracking$new(tally = rep(0,40))
for(i in 1:1000){ # simulate 100 games
  # cat("#### NEW GAME",i,"##### \n") no need to output after each iteration
  # new players for each game
  player1 <- player$new(pos = 1, verbose = FALSE, in_jail = FALSE, n_jail = 0)  
  player2 <- player$new(pos = 1, verbose = FALSE, in_jail = FALSE, n_jail = 0)
  for(i in 1:150){ # 150 turns for each game
    if(player1$verbose) cat("Player 1 turn\n")
    taketurn(player1, space_tracking)  
    if(player2$verbose) cat("Player 2 turn\n")
    taketurn(player2, space_tracking)  
  }
}
# the results after 100 turns. No rules have been implemented
library(magrittr)
library(dplyr)
results <- cbind(gameboard, tally = space_tracking$tally)
results <- cbind(results, rel = results$tally/sum(results$tally))
results <- results %>% arrange(desc(rel))
print(results)
sum(results$tally)
